---
title: "MSA_Accuracy"
output: html_document
date: "2026-03-24"
---

```{r}
library(dplyr)
library(ggplot2)
library(Hmisc)
library(viridis)
library(table1)
```

### Summary table MSA Methods

```{r}
table1(~TCS | MSAMethod, data=Summary_TC_Scores)
```


### MSA Methods Mean & sd

```{r}
overall_mean <- mean(Summary_TC_Scores$TCS)
overall_sd <- sd(Summary_TC_Scores$TCS)
overall <- Summary_TC_Scores %>%
  group_by(MSAMethod) %>%
  summarise(
    mean_TCS = mean(TCS),
    sd_TCS = sd(TCS))
overall
```

### Overall Alignment Accuracy across MSA Methods (Error - range)

```{r}
ggplot(overall, aes(x = MSAMethod, y = mean_TCS, fill = MSAMethod)) +
  geom_col(colour = "black") + 
  geom_errorbar(aes(ymin = mean_TCS - sd_TCS,
                    ymax = mean_TCS + sd_TCS),
                width = 0.2) +
  scale_fill_manual(values = c("#DF4F4F", "#CFCF24", "#2BE62B", "#3FDADA", "#3232D4", "#DD2DDD")) +
  labs(x = "MSA Method",
       y = "Alignment accuracy (Total Column Score)",
       title = "Overall Alignment Accuracy across MSA Methods") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

# uses colour blind friendly palette - from https://rgblind.com/color-blind-palette
```

### Overall Alignment Accuracy across MSA Methods (Error - standard deviation)

```{r}
range_df <- Summary_TC_Scores %>%
  group_by(MSAMethod) %>%
  summarise(
    mean_TCS = mean(TCS),
    min_TCS = min(TCS),
    max_TCS = max(TCS))

ggplot(range_df, aes(x = MSAMethod, y = mean_TCS, fill = MSAMethod)) +
  geom_col(colour = "black") +
  geom_errorbar(aes(ymin = min_TCS,
                    ymax = max_TCS),
                width = 0.2) +
  labs(
    x = "MSA Method",
    y = "Alignment accuracy (Total Column Score)",
    title = "Overall Alignment Accuracy across MSA Methods") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none")
```

### Heatmap - Alignment Accuracy across MSA Methods

```{r}
library(viridis)
ggplot(Summary_TC_Scores, aes(x = Alignment, y = MSAMethod, fill = TCS)) +
  geom_tile(colour = "black", linewidth = 0.5) +
  scale_fill_viridis(option = "turbo",
                     direction = 1,
                     name = "TC Score") +
  labs(
       x = "Condition",
       y = "MSA Method") +
  theme_minimal(base_size = 12) 
```

### Boxplot - Alignment Accuracy with MSA Methods vs. Total Column Score

```{r}
ggplot(data = Summary_TC_Scores, aes(x = MSAMethod, y = TCS)) +
  geom_boxplot(aes(color = MSAMethod), width = 0.6, size = 1, show.legend = FALSE) +
  stat_summary(fun = mean, colour = "black", geom = "point", shape = 18, size = 4) +
  scale_fill_manual(values = c("#DF4F4F", "#CFCF24", "#2BE62B", "#3FDADA", "#3232D4", "#DD2DDD")) +
  geom_jitter(aes(color = MSAMethod), 
              width = 0.1,
              size = 2,
              alpha = 0.5,
              show.legend = FALSE) +
  coord_cartesian(ylim = c(0.5, 1)) +
  theme_minimal(base_size = 12) +
  labs(x = "MSA Method", y = "Total Column Score")
```

### Lineplot - Gap Length vs. Mean TC Score 

#### (i.e. taking the Mean TC Score of Gap Length of 10 as the mean of Pattern A and Pattern B with deletions of 10 residues)

```{r}
df_gap_line <- Summary_TC_Scores %>%
  group_by(MSAMethod, GapLength) %>%
  summarise(mean_TCS = mean(TCS))

ggplot(df_gap_line, aes(x = GapLength, y = mean_TCS, group = MSAMethod, colour = MSAMethod)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(10, 30)) +
  scale_fill_manual(values = c("#DF4F4F", "#CFCF24", "#2BE62B", "#3FDADA", "#3232D4", "#DD2DDD")) +
  theme_minimal(base_size = 14) +
  coord_cartesian(ylim = c(0.75, 1)) +

  labs(
    x = "Gap Length (amino acids)",
    y = "Mean Total Column Score"
  ) +
  theme_minimal(base_size = 12)+
  theme(aspect.ratio = 0.6)
```

### Lineplot - Gap Pattern vs. Mean TC Score 

#### (i.e. taking the Mean TC Score of Gap Pattern A as the mean of deletions of 10 and 30 residues with gap pattern A)

```{r}
df_pattern_line <- Summary_TC_Scores %>%
  group_by(MSAMethod, GapPattern) %>%
  summarise(mean_TCS = mean(TCS))

ggplot(df_pattern_line, aes(x = GapPattern, 
                            y = mean_TCS, 
                            group = MSAMethod, 
                            colour = MSAMethod)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_fill_manual(values = c("#DF4F4F", "#CFCF24", "#2BE62B", "#3FDADA", "#3232D4", "#DD2DDD")) +
  theme_minimal(base_size = 14) +
  coord_cartesian(ylim = c(0.75, 1)) +

  labs(
    x = "Gap Pattern (A = overlapping, B = non-overlapping)",
    y = "Mean Total Column Score") +
  theme_minimal(base_size = 12) +
  scale_x_discrete(drop = FALSE) +
  theme(aspect.ratio = 0.6)
```

```{r}
library(ggplot2)
library(viridis)

ggplot(Summary_TC_Scores, aes(x = MSAMethod, y = TCS, colour = MSAMethod)) +
  geom_point(size = 2) +
  facet_wrap(~ paste(GapLength, GapPattern, sep = "aa gap - Pattern "), ncol = 2) +
  scale_fill_manual(values = c("#DF4F4F", "#CFCF24", "#2BE62B", "#3FDADA", "#3232D4", "#DD2DDD")) +
  labs(
    x = "MSA Method",
    y = "Total Column Score (TCS)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none")
```


